---
title: "STAT 231: Problem Set 6B"
author: "Molly Cooper"
date: "due by 5 PM on Friday, October 9"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This homework assignment is designed to help you futher ingest, practice, and expand upon the material covered in class over the past week(s).  You are encouraged to work with other students, but all code and text must be written by you, and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps6B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps6B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(wordcloud)
library(textdata)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 
# If you discussed this assignment with any of your peers, please list who here:

> ANSWER:

\newpage
# Trump Tweets

David Robinson, Chief Data Scientist at DataCamp, wrote a blog post ["Text analysis of Trump's tweets confirms he writes only the (angrier) Android half"](http://varianceexplained.org/r/trump-tweets/).

He provides a dataset with over 1,500 tweets from the account realDonaldTrump between 12/14/2015 and 8/8/2016.  We'll use this dataset to explore the tweeting behavior of realDonaldTrump during this time period.

First, read in the file. Note that there is a `TwitteR` package which provides an interface to the Twitter web API.  We'll use this R dataset David created using that package so that you don't have to set up Twitter authentication.  

```{r}
library(twitteR)
load(url("http://varianceexplained.org/files/trump_tweets_df.rda"))
```

## A little wrangling to warm-up

1a.  There are a number of variables in the dataset we won't need.  

- First, confirm that all the observations in the dataset are from the screen-name `realDonaldTrump`.  

- Then, create a new dataset called `tweets` that only includes the following variables:

- `text`
- `created`
- `statusSource`

```{r}
#Confirming all tweets are from `realDonaldTrump`
unique(trump_tweets_df$screenName)

#Including only desired variables
tweets <- trump_tweets_df[c(1, 5, 10)]

```

\newpage
1b. Using the `statusSource` variable, compute the number of tweets from each source.  How many different sources are there?  How often are each used?

> ANSWER: There are 5 different sources: (1 is used 762 times: 2 is used 628 times: 3 is used 120 times: 4 and 5 are used once).

```{r}
#Counting and sorting the number of twitter sources
tweetsource <- tweets %>%
  count(statusSource) %>%
  arrange(desc(n))

```

\newpage
1c. We're going to compare the language used between the Android and iPhone sources, so only want to keep tweets coming from those sources.  Explain what the `extract` function (from the `tidyverse` package) is doing below.  (Note that "regex" stands for "regular expression".)

> ANSWER: The extract function below is creating a new column "source" which extracts the character "Android" or "iPhone" from the status source code to more easily determine which source the tweet came from. Esentially it's creating a column which "extracts" its information from an already exisiting column. 

```{r}
tweetdata <- tweets %>%
  extract(col = statusSource, into = "source"
          , regex = "Twitter for (.*)<"
          , remove = FALSE) %>%
  filter(source %in% c("Android", "iPhone"))
```


\newpage
## How does the language of the tweets differ by source?  

2a. Create a word cloud for the top 50 words used in tweets sent from the Android.  Create a second word cloud for the top 50 words used in tweets sent from the iPhone.  How do these word clouds compare?  (Are there some common words frequently used from both sources? Are the most common words different between the sources?)

Don't forget to remove stop words before creating the word cloud.  Also remove the terms "https" and "t.co".

> ANSWER:  There are a lot of similarities between these two word clouds including words like: "hillary" , "crooked", "trump" etc... this association is not suprising at all given that most of the common words were simply the most pertinent to Trump's 2016 campaign and thus would naturally be used frequently regardless of the source when tweeting about the election. The main differentiation is that the iPhone sourced tweets use many more names of states which perhaps suggests that this was the phone Trump took with him while tweeting on the campaign trail and traveling to various states or that perhaps his tweets from the iPhone are more informational and the tweets from the Android tent to be more impassioned. 

```{r, fig.width=8, fig.height=8}
# loading the stop_word data set
data(stop_words)
head(stop_words)
tail(stop_words)
stop_words %>% count(lexicon)

#transforming text to word
texttoword <- tweetdata %>%
  unnest_tokens(output = word, input = text) 

#adding more stop words
morestopwords <- tribble(
  ~word, ~lexicon,
  "https", "CUSTOM",
  "t.co", "CUSTOM"
)

#joining all of our stop words
stop.words <- stop_words  %>% bind_rows(morestopwords)

#stop words removed
commonwords <- texttoword %>%
  anti_join(stop.words , by="word")
```


```{r}

# Android word cloud

pal <- brewer.pal(10, "Paired")

AndroidWords <- commonwords %>% 
  filter(source == "Android")%>%
  count(word)

wordcloud(
  words = AndroidWords$word,
  freq = AndroidWords$n,
  max.words = 50,
  scale=c(2,0.5),
  colors = pal
)

# iPhone word cloud

iPhoneWords <- commonwords %>% 
  filter(source == "iPhone")%>%
  count(word)

wordcloud(
  words = iPhoneWords$word,
  freq = iPhoneWords$n,
  max.words = 50,
  scale=c(2,0.5),
  colors = pal
)
```



\newpage
2b. Create a visualization that compares the top 10 *bigrams* appearing in tweets by each source (that is, facet by source).  After creating a dataset with one row per bigram, you should remove any rows that contain a stop word within the bigram.  

How do the top used bigrams compare between the two sources?

> ANSWER: The top used bigrams between the two sources demonstrate how different the tone of the language used in Trump's iPhone and Android tweets is: "join me", "makeamericagreatagain", and "thank you" are frequently used in his iPhone tweets which correspond to formal language and his campaign slogans whereas the most frequently used rhetoric in his Android tweets is: "crooked hillary", "bernie sanders", and "ted cruz" which suggests that he's typically using this source in a more informal manner to trash-talk other politicians rather than focusing on his own campaign. 


```{r}
#Cutting stop words that were shown in the original visualization
stopwords <- tribble(
  ~word, ~lexicon,
  "https t.co", "CUSTOM",
  "will be", "CUSTOM",
  "trump2016 https", "CUSTOM",
  "of the", "CUSTOM",
  "makeamericagreatagain https", "CUSTOM",
  "is a", "CUSTOM",
  "in the", "CUSTOM",
  "i will", "CUSTOM",
  "for the", "CUSTOM",
  "to the", "CUSTOM",
  "americafirst https", "CUSTOM",
  "you for", "CUSTOM",
  "i am", "CUSTOM",
  "me in", "CUSTOM",
  "to be", "CUSTOM"
)

#Creating the bigram data
visual <- tweetdata %>% 
  unnest_tokens(output = word, input = text,token="ngrams",n=2) %>% 
  anti_join(stopwords, by = "word") %>%
  group_by(source) %>%
  count(word, sort = TRUE) %>% 
  group_by(source) %>%
  slice(1:10)
  
#Creating a visualization for the bigram dataset
  ggplot(data = visual, mapping = aes(x = word, y = n, color = word, fill=word)) +
  geom_col(show.legend =FALSE) +
  xlab(NULL) +
  ylab("Word Frequency in Trump's Tweets") +
  coord_flip() +
  facet_wrap(~source)
```


\newpage
2c. Consider the sentiment.  Compute the proportion of words among the tweets within each source classified as "angry" and the proportion of words classified as "joy"  based on the NRC lexicon.  How does the proportion of "angry" and "joy" words compare between the two sources?  What about "positive" and "negative" words?  

> ANSWER: 

> Anger -> (Android):363/2447=0.15 , Anger(iPhone):170/2165=0.8 
> Joy -> (Android):292/2447=0.12 , (iPhone):176/2165=0.8
> Positive -> (Android):794/2447=0.32 , (iPhone):449/2165=0.21
> Negative -> (Android):662/2447=0.27, (iPhone): 266/2165=0.12

> Based on my calculations it seems that Trump uses more "Angry" and "Negative" words on his Adroid but also more "Joy" and "Positive" words. I would extrapolate that Trump is sending impassioned and personal opinions on his Android and he is sending more neutral and informative tweets about his campaign from his iPhone. 

```{r}
nrc_lexicon <- get_sentiments("nrc")

#Finding the number of words from each source that correspond to each emotion
emotion <- commonwords%>%
  inner_join(nrc_lexicon, by = "word") %>% 
  group_by(source) %>%
  count(sentiment)

#Total number of words
total <- emotion$n

#Out of curiosity I calculated the overall proportion of his tweets by sentiment
emotion <- emotion%>%
  mutate(prop = n/sum(total))

emotion



```


\newpage
2d. Lastly, based on your responses above, do you think there is evidence to support Robinson's claim that Trump only writes the (angrier) Android half of the tweets from realDonaldTrump?  In 2-4 sentences, please explain.

> ANSWER: All of my responses do support this theory. To expand on just the emotion behind his tweets it seems that Trump is perhaps using his iPhone in a more professional manner and his Android in a personal manner. This likely indicates that a campaign manager or coordinator is controlling and/or writing the tweets that are sent from his iPhone to inform and encourage his public image. While his Android tweets are likely written by himself and used as a platform for his personal opinions and frusturations. It's very interesting how much insight we can glean from analyzing language through R and pinpointing the differences between seemingly unimportant distinctions from a source code. 

